Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-04 paper/main.tex
Updates

\subsection{Using LLMs as Experts}
Recently, there has been a lot of work towards utilizing LLMs for causal
discovery including approaches ranging from determining pairwise edge
orientation \citep{Kiciman2023, Jin2024} to full causal discovery
\citep{Naik2023, Vashishtha2023}, and counterfactual reasoning
tasks\citep{Kiciman2023} (see \citet{Liu2024} for an overview) As our approach
utilizes expert knowledge for determining the orientation of the edges, this
opened up an opportunity to use an LLM for determining this. We tried applying
our procedure for causal discovery with using GEMINI as the expert. We used a
greedy approach where the pair of variables with the highest unexplained
correlation is oriented first. The prompt used to determine the orientation is
provided in Appendix~\ref{section:llms}. Our LLM prompt utilizes variable
description to determine the edge orientation. We applied this approach to the
adult income dataset and the output is shown in Fig.~\ref{fig:adult_llm}.

>>>>

\subsection{Using LLMs as Experts}
Recently, there has been a lot of work towards using LLMs for causal discovery,
with tasks ranging from determining pairwise edge orientation
\citep{Kiciman2023, Jin2024} to full causal structure learning \citep{Naik2023,
Vashishtha2023} and counterfactual reasoning tasks\citep{Kiciman2023} (see
\citet{Liu2024} for an overview).

Since our approach utilizes expert knowledge to determine edge orientations, we
investigated the potential of using an LLM for this task. We applied our causal
discovery procedure with using Gemini 1.5 flash model as the expert. Our method
uses a greedy approach where the pair of variables with the highest unexplained
correlation is oriented first. The prompt used to determine the orientation is
provided in Appendix~\ref{section:llms}. Our LLM prompt utilizes variable
description to determine the edge orientation. We applied this approach to the
adult income dataset and the output is shown in Fig.~\ref{fig:adult_llm}.

Unlike the other approaches to utilize LLM for doing full causal discovery, we 
used a combination of statistical tests and LLMs. This greatly reduces the
amount of information that we need from the LLM. Our approach is also able to 
ask exact questions about the variables unlike some of the other approaches.
